snake change

### 2023-03-09 graphs_newt_snakes/Paper_figs.Rmd
I have been filling out my thoughts about these experiments. They are truly some cool results

@@

@@

GA1_cor<- read.csv("~/Desktop/GA1_tall_redo/all_cor.txt" , header = TRUE)
GA1_cor$group_rep <- paste0(GA1_cor$snake_mu_rate, "_", GA1_cor$snake_mu_effect_sd,"_",GA1_cor$newt_mu_rate, "_", GA1_cor$newt_mu_effect_sd, "_", GA1_cor$rep)
GA1_cor$group <- paste0(GA1_cor$snake_mu_rate, "_", GA1_cor$snake_mu_effect_sd,"_",GA1_cor$newt_mu_rate, "_", GA1_cor$newt_mu_effect_sd)
GA1_cor$sim=LETTERS[factor(GA1_cor$group)]
GA1_cor$exp="Flat Map"

@@

GA1_lit<- read.csv("~/Desktop/GA1_tall_redo/all_lit.txt" , header = TRUE)
GA1_lit$group_rep <- paste0(GA1_lit$snake_mu_rate, "_", GA1_lit$snake_mu_effect_sd,"_",GA1_lit$newt_mu_rate, "_", GA1_lit$newt_mu_effect_sd, "_", GA1_lit$rep)

>>>>

GA1_final_csv<- read.csv("~/Desktop/GA1_final_csv/all_grid.txt" , header = TRUE) %>%  filter(between(generation, 40000, 50000))
GA1_final_csv$group_rep <- paste0(GA1_final_csv$snake_mu_rate, "_", GA1_final_csv$snake_mu_effect_sd,"_",GA1_final_csv$newt_mu_rate, "_", GA1_final_csv$newt_mu_effect_sd, "_", GA1_final_csv$rep)
GA1_final_csv$group <- paste0(GA1_final_csv$snake_mu_rate, "_", GA1_final_csv$snake_mu_effect_sd,"_",GA1_final_csv$newt_mu_rate, "_", GA1_final_csv$newt_mu_effect_sd)
GA1_final_csv$sim=LETTERS[factor(GA1_final_csv$group)]
GA1_final_csv$grid_num = 5

@@

file=GA1_final_csv
file_name="1e-11_5_1e-10_0.5_0"
spegen=44501
type="mean"
row_num=20
col_num=5

@@

GA1_cor<- read.csv("~/Desktop/GA1_final_csv/all_cor.txt" , header = TRUE)
GA1_cor$group_rep <- paste0(GA1_cor$snake_mu_rate, "_", GA1_cor$snake_mu_effect_sd,"_",GA1_cor$newt_mu_rate, "_", GA1_cor$newt_mu_effect_sd, "_", GA1_cor$rep)
GA1_cor$group <- paste0(GA1_cor$snake_mu_rate, "_", GA1_cor$snake_mu_effect_sd,"_",GA1_cor$newt_mu_rate, "_", GA1_cor$newt_mu_effect_sd)
GA1_cor$sim=LETTERS[factor(GA1_cor$group)]
GA1_cor$exp="Flat Map"

@@

GA1_lit<- read.csv("~/Desktop/GA1_final_csv/all_lit.txt" , header = TRUE)
GA1_lit$group_rep <- paste0(GA1_lit$snake_mu_rate, "_", GA1_lit$snake_mu_effect_sd,"_",GA1_lit$newt_mu_rate, "_", GA1_lit$newt_mu_effect_sd, "_", GA1_lit$rep)

### 2023-03-09 graphs_newt_snakes/dif_cost_grade.Rmd
I have been filling out my thoughts about these experiments. They are truly some cool results

@@

What does it mean to be correlated?

@@

@@

Each GA combination and trial has its own msprime simulation, but the msprime file is shared between the 5 section and 7 section slim runs.

### The data

@@

@@

@@

@@

whale_gen(cost_grade_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

@@

@@

### cor
```{r, combo myplots}

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

The goal of this experiment was to create a gradient picture and to assign different cost values to areas of the map

@@

What does it mean to be correlated? How might an uneven environment (its easier to have a higher phenotype in some area of the map) impact co-evolution?

@@

Previously I have been looking at how newts and snakes can co-evolve under different genetic architectures (mutation rate and mutation effect size), and have found that while the mean phenotype increases there is no spatial phenotype correlation. This has puzzled me, how can there be a constant increase in phenotype but no spatial phenotype correlation? I think that there might be other factors that lead to spatial phenotype correlation.

@@

Each GA combination and trial has its own msprime simulation.

I created a gradient map where at the top there is a higher fitness cost for having a large phenotype (set to 50) and at the bottom there is a lower cost for having a higher phenotype (set to 250). Both newts and snakes are effected by this gradient.

![Gradient Map](/Users/victoria/Desktop/Coevo/newt_snake/data/gradent.png)

@@

In this first section I look at the entire populations mean phenotype for both snakes (blue) and newts (red). The difference between mean snake and mean newt phenotype is shown on the black line. For each GA combination there are 4 sets of lines (red, blue, black). Each line is a different trial with the same simulation parameters. I also present the average difference (of the trials) between snake and newt phenotype in the table of average differences.

@@

These results look very similar to the results seen in my other simulations where there is no gradient map. The mean phenotypes of newts and snakes go up, until they reach an equilibrium. Here the aver values are higher than what I have seen before.

@@

Here I plot the interaction between newt/snake phenotype and population size. Typically, when a species had a higher phenotype they also had a larger population size. This relation between phenotype and population size had specific outcomes that depended on the GA of newts and snakes.

The first figure compares the population size of newts and snakes to the difference between mean snake and mean newt phenotype for a time slice (5,000-10,000 generations). Color in this plot is the difference between snake and newt phenotype, with blue indicating snakes have a larger phenotype and red indicating newts have a larger phenotype. Cream color points indicate that the two phenotypes are nearly the same. The second figure present the histograms of the difference between snake and newt population size (green) and phenotype (purple) for a time slice (5,000-10,000 generations).

@@

whale_gen(cost_grade_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

@@

These results are similar to what I have seen before. There might be an increase in the differences between snake and newt phenotype, but the overall message higher phenotype = more individuals in unchanged. GA3 seems to create the larger difference in mean newt/snake phenotype.

@@

The next section I am examining the spatial correlation between newt and snake phenotypes and I predicted that there would be a positive correlation between the phenotypes. I first look at the correlation between mean newt phenotype and mean snake phenotype for each of the four trials in every GA combination from 10,000-15,000 generations. The solid line is a 0 with a dashed line at the level of correlation seen in natural newt-snake population(s).

```{r, combo myplots}

@@

In my previous results (no gradient map) I saw no positive spatial phenotype correlation between snakes and newts. Here, with a cost gradient map I see that there are quite a few simulations where newt and snake local mean phenotypes are correlated when compared across space.

@@

In order to understand how spatial correlations where changing with time I took 5,000 generation time slices to look at all four trials correlation values. Each color is a different trial per GA combination. The histogram values are stacked.

@@

Spatial phenotype correlation increases for most of the simulation as I move through the time slices. Once a simulation reach a high spatial phenotype correlation, it did not go back down. The simulations where non surpassed the empirical newt and snake spatial phenotype correlation all contained a genetic architecture that had a small mutational varance.

@@

Next, I examine three randomly chosen plots. Time (in generations) in on the x-axis and both mean phenotype and phenotype spatial correlation in on the y-axis. Newt whole population mean phenotype is red, while snake mean phenotype is blue. The pink line is the phenotype spatial correlation.

@@

Most of these plots show that as time increases both the mean newt & snake phenotype as well as the   spatial phenotype correlation (pink line) increase (very few show no change). Spatial phenotype correlation increases the most when the newt and snake mean phenotypes are increasing the most. Some of these plots are very interesting to see.

@@

This next section is just getting a glimpse at how newt & snake phenotype and population size differ over time. The populations start off with about 250 individuals each. Each individual has a different genetic background created from msprime. Then each msprime simulation is put into slim and data is generated. Plots show newt by snake population size, with the point color representing the difference between mean snake and newt phenotype (red=newts have a higher phenotype and blue=snakes have a higher phenotype). The other plots show histograms of difference between snakes and newts phenotype and population size (purple and green).

@@

The results looked similar to results that I have seen in other experiments. In the beginning of the simulation both newt and snake population grows. The difference in phenotype quickly becomes polarized. The population size reaches a steady point and then newts and snakes co-evolve. In the middle part of my simulation, the difference between newt and snakes mean phenotype solidifies. Often, the difference in mean phenotype decreases (where compared to the beginning of the simulation). When the GA has a high mutation rate and low mutation effect size (GA 1), the difference in mean phenotype grows. This leads to the species with GA 1 losing the co-evolutionary arms race. More frequent smaller steps does not help a species win in an arms race (might also be due to lower mutational variance). The histograms reflect what is seen in the scatter plots.

@@

In the summary section, I try to come up with a way to show how different GA combinations can change the simulations results. In all of these plots snakes GA is represented by color and newt GA is represented by shape. There 16 color-shape combinations (with 4 repeats for trials). There are four sets of plots: 1) newt by snake population size, 2) phenotype difference by snake population size, 3) phenotype difference by snake GA, and 4) phenotype difference by newt GA. There are three figures in each set, taken at the begging, middle, and end time chunks.

@@

Firstly, these plots are very similar to the plots seen without a gradient map and have the same results. There are clusters of points that create lines of shapes and lines of colors. The best GA for newts and snakes was 1e-09_0.05 (can be seen as green near the top of all the lines of shapes and triangles near the bottom right look at all the lines of color). It is interesting to see how these points spread apart, but remain similar between trials. When putting these figures together it seems like the population size of snakes is lower when the newt phenotype is larger than the snake phenotype.

@@

In the heatmap plots each GA combination and trails is presented by combining newt GA in the x-axis to snake GA and trial number in the y-axis. The result is the color in that section. There are two types of heatmap plots shown below. One shows the average snake population size for a time chunk with darker colors indicating a smaller snake population and lighter colors indicating a larger snake population. The other heatmap shows the average difference between snake and newt mean phenotype (red=newts had a higher phenotype, blue=snakes had a higher phenotype). I look at 3 time slices for both types of heatmaps.

@@

Heatmap results are very similar to the no cost gradient maps. They star off sort of neutral, but each trial of the GA combination creates very similar patterns. Which shows that under certain GA either newts or snakes have an co-evolutionary advantage (1e-9(0.05)^2 = 2.5e-12 or 1e-10(0.5)^2 = 2.5e-11). The cost gradient results look more dramatic than the results with out the map.

@@

This section goes over the results from the local measurements (grid calculations). I divided my map up into smaller area (grids) and calculated mean phenotype, max phenotype, min phenotype, and population size. In each of these plots newts are represented by circles and snakes are represented by squares. Parameter values increase from a dark color to a lighter color (green-blue themed for phenotype, orange-pinked themed for population size) There is also a subplot that plots each parameter (mean, max, …) of newt by snake colored by map location (red=corner, green=edge, blue=middle). I look at the one simulation at one time in the begging and end.

@@

Newt and snake phenotypes are positively spatially correlated. Newt and snake population sizes are negativly correlated (look at the mini plots and not the values)

### 2023-03-09 graphs_newt_snakes/interrate_grade.Rmd
I have been filling out my thoughts about these experiments. They are truly some cool results

@@

@@

Each GA combination and trial has its own msprime simulation, but the msprime file is shared between the 5 section and 7 section slim runs.

### The data

@@

@@

whale_gen(interrate_grade_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

The goal of this experiment was to create a gradient picture and to assign different interaction values to areas of the map

@@

Previously I have been looking at how newts and snakes can co-evolve under different genetic architectures (mutation rate and mutation effect size), and have found that while the mean phenotype increases there is no spatial phenotype correlation. After adding a cost gradient map and seeing high phenotype spatial correlations I wondered if there were other things that could vary and still create spatial phenotype correlation.

@@

Each GA combination and trial has its own msprime simulation. I created a gradient map where at the top there is a lower rate of newt and snake interaction (0.01) and at the bottom there is a higher rate of newt/snake interaction (0.1 so high that there might be local species extinction). My simulation typically use 0.05 as an interaction rate (interaction rate was explores in the interaction rate simulations)

@@

These results follow a similar pattern to what I have seen in other simulations. Both newt and snake mean phenotype go up, but they go up slowly and reach a low equilibrium. The traits do not seem exaggerated, but they do not head towards zero. In simulations with a low mutational variance GA one of the species mean phenotype flat line.  I wonder if the results would change if the interaction rate was different.

@@

whale_gen(interrate_grade_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

@@

The overall meaning of these plots suggest that a higher phenotype leads to more individuals. The dot smear is longer than what I have seen in my other simulations. I wonder if the varying the interaction rate leads to more stocatic values.

@@

The spatial phenotype correlations of this simulation set are positive (when excluding low mutational variance GA sims). Many of the spatial phenotype correlation values are near the real newt/snake phenotype correlation.

@@

These plots reveal that the spatial phenotype correlation between snakes and newts increase with time and end up being close to the empirical newt snake spatial phenotype correlation. These correlation values are higher than when there is no gradient map, but lower then the cost gradient map. There is no spatial phenotype correlation when the GA has low mutational variance.

@@

Most of these plots show that as time increases both the mean newt & snake phenotype (read and blue lines) as well as the spatial phenotype correlation (pink line) increase (very few show no change). Spatial phenotype correlation increases the gradually as the newt and snake mean phenotypes gradually increases. There is a lot of variation of the spatial phenotype correlation between different time points (every 20 generations) (it seems to go up and down quiet a bit).

@@

his next section is just getting a glimpse at how newt & snake phenotype and population size differ over time. The populations start off with about 250 individuals each. Each individual has a different genetic background created from msprime. Then each msprime simulation is put into slim and data is generated. Plots show newt by snake population size, with the point color representing the difference between mean snake and newt phenotype (red=newts have a higher phenotype and blue=snakes have a higher phenotype). The other plots show histograms of difference between snakes and newts phenotype and population size (purple and green).

@@

In the beginning of the simulation both newt and snake population grows. The difference in phenotype quickly becomes polarized. The population size reaches a steady point as the newts and snakes co-evolve. In this experiment set the point clusters are more elongated, which suggests that the newt and snake population size flucates more than what I have seen in my other experiments. When the GA has a high mutation rate and low mutation effect size (GA 1), the difference in mean phenotype grows. This leads to the species with GA 1 losing the co-evolutionary arms race. The histograms reflect what is seen in the scatter plots.

@@

n the summary section, I try to come up with a way to show how different GA combinations can change the simulations results. In all of these plots snakes GA is represented by color and newt GA is represented by shape. There 16 color-shape combinations (with 4 repeats for trials). There are four sets of plots: 1) newt by snake population size, 2) phenotype difference by snake population size, 3) phenotype difference by snake GA, and 4) phenotype difference by newt GA. There are three figures in each set, taken at the begging, middle, and end time chunks.

@@

These plots mimic the plots seen with and without a cost gradient map, where clusters of points create lines of shapes and lines of colors, but there are some differences. In the late newt popsize by snake popsize the points seem to be spread out more (like more extreme population sizes). In the late phenotype difference and snake population size there is more of a difference between snake population size and the difference between snake and newt mean phenotype (most other plots things are more linear). For GA 1e-11_5 (purple points for snakes and plus points for newts) seem to have lead to a lower phenotype (or larger phenotype difference, i.e. when snakes have this GA newts tend to have a higher mean phenotype). These results puzzles me.

@@

In the heat maps there is more variance between in both snake population size and phenotype difference simulation trials. The results tell similar stories when the GA of a species is 1e-09_0.05 the population size and phenotype are higher. Something is going on that is changing the difference between snake and newt phenotype in simulations with a GA 1e-11_5. These phenotype results are similar to the results seen when a species has a GA 1e-09_0.05. Something about a gradient interaction rate is changing the phenotype results. I think it might be the way things are averaged across the entire population. For example, if mutation that have a high cost lead to the individual to not reproducing, we lose out on higher phenotypes (for GA=1e-11_5) which we would see in areas of higher interaction. If we average between areas with few interactions (low phenotype) at the top and more interactions at the bottom (higher phenotype), but there are less individuals with a higher phenotype due to cost, the over all population average would be smaller.

@@

In the beginning of this particular simulations results look very spatially correlated, but later in this simulation there are not enough newts at the bottom of the map to make a good spacial correlation calculation so results look uncorrelated.

### 2023-03-09 graphs_newt_snakes/newt_change.Rmd
I have been filling out my thoughts about these experiments. They are truly some cool results

@@

@@

Each GA combination and trial has its own msprime simulation, but the msprime file is shared between the 5 section and 7 section slim runs.

### The data

@@

@@

whale_gen(newt_change_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

@@

@@

@@

@@

@@

@@

@@

>>>>

To test what would happen with the spatial phenotype correlation between newts and snakes when the fitness cost for newts in on a scale (dependent on location) and the fitness cost for snakes is constant.

@@

Previously I have been looking at how newts and snakes can co-evolve under different genetic architectures (mutation rate and mutation effect size), and have found that while the mean phenotype increases there is no spatial phenotype correlation. This has puzzled me, how can there be a constant increase in phenotype but no spatial phenotype correlation? I think that there might be other factors that lead to spatial phenotype correlation. When I used a cost gradient map, I found that newt and snakes mean phenotype were spatially correlated. This is similar to an environmental condition. But what if the condition was only applies to one speceices?

@@

Each GA combination and trial has its own msprime simulation. I created a gradient map where at the top there is a higher fitness cost for having a large phenotype and at the bottom there is a lower cost for having a higher phenotype. Only newts are effected by this gradient. Snakes fitness cost remain at 100. To see the gradient figure see cost gradient file.

@@

These results look very similar to the results seen in my other simulations (no gradient map and gradient map). The mean phenotypes of newts and snakes go up, until they reach an equilibrium. Here the average values are higher than what I have seen in the no gradent map, but are lower than what I see in the cost gradient map.

@@

whale_gen(newt_change_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

@@

These results are similar to what I have seen before. There might be an increase in the differences between snake and newt phenotype, but the overall message higher phenotype = more individuals in unchanged.

@@

In my previous results (no gradient map) I saw no positive spatial phenotype correlation between snakes and newts. Here, with a cost gradient map for newts I see that there are quite a few simulations where newt and snake local mean phenotypes are correlated when compared across space. This correlation is not as high or as frequent as what I saw when both newt and snake fitness cost was dependent on location.

@@

Spatial phenotype correlation increases for most of the simulations as time increases. Once a simulation reach a high spatial phenotype correlation, it typically did not go back down. The simulations where the simulations correlation values did not come close to the empirical newt and snake spatial phenotype correlation all contained a genetic architecture that had a small mutational variance. The exception to this is in the simulation where both newts and snakes had a low mutational variance, the local phenotypes became correlated then it faded (I wonder why). The spatial phenotype correlation is a lot higher than when there was not gradient map, but lower than when both newt and snakes fitness cost was on a gradient. When compared to the correlation results of snakes on a cost gradient and newts constant, the results in this document are more spatially correlated.

@@

The results looked similar to results that I have seen in other experiments. In the beginning of the simulation both newt and snake population grows. The difference in phenotype quickly becomes polarized. The population size reaches a steady point and then newts and snakes co-evolve. In the middle part of my simulation, the difference between newt and snakes mean phenotype solidifies. Often, the difference in mean phenotype decreases (where compared to the beginning of the simulation). When newts had a GA that with a high mutation rate and low mutation effect size (GA 1), the difference in mean phenotype grows, leading to the newts losing the co-evolutionary arms race. In all other GA combinations newts had a larger phenotype. When looking at the snakes on a cost fitness scale, the results are flipped. When snakes had a GA that with a high mutation rate and low mutation effect size (GA 1), leading to the snakes losing the co-evolutionary arms race. Could there be something with in the environments causing one species to out compete the other? The histograms reflect what is seen in the scatter plots.

@@

Firstly, these plots are very similar to the plots seen with and without a gradient map. There are clusters of points that create lines of shapes and lines of colors. Points began clustered then spread apart. These points are similar between trials. There is one striking differences in these results, it seems that the results are shifted toward the newts favor (in the snakes cost gradient results things are shifted towards the snakes side). The shift does not seem to effect the circles (newt GA1). Thes plots are fun to look at.

@@

Heatmap results are different, the seem more muted (less differences between GA combinations). They start off sort of neutral, and then become set in the favor of the newt (snake has a smaller population size while newts have a larger phenotype). Each trial of the GA combination creates very similar patterns.

@@

Newt and snake phenotypes are positively spatially correlated. Newt and snake population sizes are negativly correlated (look at the mini plots and not the values). Phenotypes in this simulation did not get as high as seen in the cost gradient simulations.

### 2023-03-09 graphs_newt_snakes/no_inter_cost_grade.Rmd
I have been filling out my thoughts about these experiments. They are truly some cool results

@@

@@

Each GA combination and trial has its own msprime simulation, but the msprime file is shared between the 5 section and 7 section slim runs.

### The data

@@

### Phenotype differences 
```{r, graphing by gen}
trends_by_gen(no_interaction_cost_grade_lit, title="Space Tall: Newt VS Snake Population Size GA1", subt="Gen 16sims", y_label="Mean Phenotype", x_label="Generations", 100, 50000, rep=1)

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

The goal of this experiment was to create a gradient picture and to assign different cost values to areas of the map and turn the interaction off. This will see if the interaction is needed to create spatial phenotype correlations between newts and snakes.

@@

Previously I have been looking at how newts and snakes can co-evolve under different genetic architectures (mutation rate and mutation effect size), and have found that while the mean phenotype increases there is no spatial phenotype correlation. This has puzzled me, so I added in an environment factor (cost gradient) and found that newt and snake mean phenotypes became spatially correlated.

@@

I created a gradient map where at the top there is a higher fitness cost for having a large phenotype and at the bottom there is a lower cost for having a higher phenotype. Both newts and snakes are effected by this gradient. However, in this simulation I break the interaction between newts and snakes. Survival of the newt or snake depends on a coin flip and not their phenotype.

@@

In this first section I look at the entire populations mean phenotype for both snakes (blue) and newts (red). The difference between mean snake and mean newt phenotype is shown on the black line. For each GA combination there are 4 sets of lines (red, blue, black). Each line is a different trial with the same simulation parameters. I also present the average difference (of the trials) between snake and newt phenotype in the table of average differences.

### Phenotype differences 
```{r, graphing by gen}
trends_by_gen(no_interaction_cost_grade_lit, title="Space Tall: Newt VS Snake Population Size GA1", subt="Gen 16sims", y_label="Mean Phenotype", x_label="Generations", 1000, 50000, rep=1)

@@

These results look very similar to the results seen in the simulation with no interaction and no gradient map. The mean phenotypes of newts and snakes go down, until they reach an equilibrium.

@@

These results are similar to what I have seen in the initial no interaction simulation. Higher phenotype does not equal more individuals, a lower GA mutational variance ends up having a higher phenotype.

@@

There is no positive or negative spatail correlation. The spatial phenotype correlation is close to zero.

@@

The spatial phenotype correlation does not increase over time, it stays around zero (similar to the no interaction/cost gradient map).

@@

The mean phenotypes of newt (red) and snake (blue) go down as the number of generations increase. The spatial correlation between newt and snake mean phenotype (pink) is randomly increasing and decreasing, but hanging out around zero.

@@

The results looked similar to results that I have seen in the other no interaction experiment. In the beginning of the simulation both newt and snake population grows. The difference in phenotype quickly becomes polarized (depending on GA). The population size reaches a steady point and then newts and snakes co-evolve. In the middle part of my simulation, the difference between newt and snakes mean phenotype solidifies. The histograms reflect what is seen in the scatter plots.

@@

After examining these plots I found that there is no relationship between GA and population size or population size and phenotype. A higher newt mutational variance led to a lower newt phenotype (possibly pushed down due to cost and no interaction). These results are very similar to the no interaction/ no coat gradient simulation. It seems like the gradient map has little effect when the interaction of newt and snake is not dependent on phenotype. What if there was an interaction, but the rate was smaller?

@@

The results of this section how that there is no relationship between GA and population size (colors seem random). There is relationship between GA and phenotype. Again, the difference in phenotype is symmetrical across the diagonal (right diagonal this time). Species have a higher phenotype is they have a GA with lower mutational variance. This might be due to the mutational effect size distribution and phenotype cost.

@@

Prototypes remain low so there is no spatial correlation.

### 2023-03-09 graphs_newt_snakes/snake_change.Rmd
I have been filling out my thoughts about these experiments. They are truly some cool results

@@

@@

Each GA combination and trial has its own msprime simulation, but the msprime file is shared between the 5 section and 7 section slim runs.

### The data

@@

@@

whale_gen(snake_change_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

@@

@@

@@

@@

@@

@@

>>>>

To test what would happen with the spatial phenotype correlation between newts and snakes when the fitness cost for snakes in on a scale (dependent on location) and the fitness cost for newts is constant.

@@

Previously I have been looking at how newts and snakes can co-evolve under different genetic architectures (mutation rate and mutation effect size), and have found that while the mean phenotype increases there is no spatial phenotype correlation. This has puzzled me, how can there be a constant increase in phenotype but no spatial phenotype correlation? I think that there might be other factors that lead to spatial phenotype correlation. When I used a cost gradient map, I found that newt and snakes mean phenotype were spatially correlated. This is similar to an environmental condition. But what if the condition was only applies to one species?

@@

Each GA combination and trial has its own msprime simulation. I created a gradient map where at the top there is a higher fitness cost for having a large phenotype and at the bottom there is a lower cost for having a higher phenotype. Only snakes are effected by this gradient. Newts fitness cost remain at 100. To see the gradient figure see cost gradient file.

@@

These results look very similar to the results seen in my other simulations (no gradient map and gradient map). The mean phenotypes of newts and snakes go up, until they reach an equilibrium. Here the average values are higher than what I have seen in the no gradient map, but are lower than what I see in the cost gradient map. There are more simulations where snakes have a higher mean phenotype.

@@

whale_gen(snake_change_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

@@

The next section I am examining the spatial correlation between newt and snake phenotypes and I predicted that there would be a positive correlation between the phenotypes. I first look at the correlation between mean newt phenotype and mean snake phenotype for each of the four trials in every GA combination from 10,000-15,000 generations. The solid line is a 0 with a dashed line at the level of correlation seen in natural newt-snake population(s).

@@

In my previous results (no gradient map) I saw no positive spatial phenotype correlation between snakes and newts. Here, with a cost gradient map for snakes I see that there are quite a few simulations where newt and snake local mean phenotypes are correlated when compared across space. This correlation is not as high or as frequent as what I saw when both newt and snake fitness cost was dependent on location. It is more similar to what I saw in the newts cost gradient results.

@@

Spatial phenotype correlation increases for most of the simulations as time increases. Once a simulation reach a high spatial phenotype correlation, it typically did not go back down. The simulations where the simulations correlation values did not come close to the empirical newt and snake spatial phenotype correlation all contained a genetic architecture that had a small mutational variance. The spatial phenotype correlation is a lot higher here, than when there was not gradient map, but lower than when both newt and snakes fitness cost was on a gradient. Results presented here are less spatially correlated, when compared to the correlation results of newts on a cost gradient and snakes constant.

@@

The results looked similar to results that I have seen in other experiments. In the beginning of the simulation both newt and snake population grows. The difference in phenotype quickly becomes polarized. The population size reaches a steady point and then newts and snakes co-evolve. In the middle part of my simulation, the difference between newt and snakes mean phenotype solidifies. Often, the difference in mean phenotype decreases (where compared to the beginning of the simulation). When snakes had a GA that with a high mutation rate and low mutation effect size (GA 1), the difference in mean phenotype grows, leading to the snakes losing the co-evolutionary arms race. In all other GA combinations snakes had a larger phenotype. When looking at the newts on a cost fitness scale, the results are flipped. When newts had a GA that with a high mutation rate and low mutation effect size (GA 1), leading to the newts losing the co-evolutionary arms race. Could there be something with in the environments causing one species to out compete the other? This was not a complete symmetric image, what could make things slightly different? The histograms reflect what is seen in the scatter plots.

@@

These plots are very similar to the plots seen with and without a gradient map. There are clusters of points that create lines of shapes and lines of colors. Points began clustered then spread apart as time increases. These points are similar between trials. There is one striking differences in these results, it seems that the results are shifted toward the snakes favor (in the newt cost gradient results things are shifted towards the newt side). The shift does not seem to effect the pink points (snake GA1). These plots are fun to look at and compare.

@@

Heatmap results are different, the seem more muted (less differences between GA combinations). They start off sort of neutral, and then become set in the favor of the snake (newts have a smaller population size while snakes have a larger phenotype). Each trial of the GA combination creates very similar patterns. The differences between GA combinations are more pronounced than when newts had a fitness cost gradient, but less pronounced when compared to both newts and snakes being on a cost gradient.
